---
title: "Logistic Regression"
author: "Stefanie Molin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r template-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load the case study library
library(modelAnalyzeR)
library(dplyr)

# all data will be loaded in the parent document and referenced within each of the child sections
```

### Logisitic Regression
How it works in detail. Parameters experimented with and what they do.

*Note that this method will be performed in Python using Scikit-learn.*

```{r format_data_to_pass_to_python, echo=FALSE, eval=FALSE}
python_AS_data <- prep_AS_for_python(AS_data, days = 25)
python_TS_data <- prep_TS_for_python(TS_data, days = 25)
python_RexT_data <- prep_RexT_for_python(RexT_data, days = 30)

# write to files for transfer
write.csv(x = python_AS_data, file = params$python_AS_file, row.names = FALSE)
write.csv(x = python_TS_data, file = params$python_TS_file, row.names = FALSE)
write.csv(x = python_RexT_data, file = params$python_RexT_file, row.names = FALSE)
```

#### Results
Evaluate best performance (tuned parameters) against the criteria defined in the main document for each of the time periods.

**TO DO** Tables go here
One per AS/TS/Exec

#### ROC Curve
Graph all points on the graph as lines per AS/TS/Exec and as an overall line

#### Alogrithm Complexity  


#### Conclusions
Observations noticed on the results. Strengths/Weaknesses of the model. Any patterns?

